Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble

### 2022-08-22 docs/paper/paper.Rmd
use data up to the 15th of August

  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 19th of July 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
  We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
  Data on notified weekly cases was originally sourced from the Johns Hopkins University curated data repository (JHU, [@REF]).
  We used the latest available observed data. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
  The follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of data revisions.
  Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.

@@

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations). In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model). Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model. Of the remaining models submitted `r length(local_models)` were submitted in only one location. These models were gernerally submitted to a subset of locations with particular clustering in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project). Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else. `r nrow(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK. We first derived a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs and our observations of the performance of other models. We restricted ourselves to a model that could be easily understand, that produced meaningful summary statistics, and that could be run with low compute resources. We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.

Over the 6 months of the study period, we found that our surrogate model performed adequately at reproducing the forecast hub ensemble with average performance that was in general slightly worse though this varied by location and forecast date. Performance was closest for short forecast horizons and degraded relative to the forecast hub ensemble as the horizon increased. We observed that our surrogate model displayed similar levels of robustness to the forecast ensemble with few outlier forecasts unlike our previous forecasting efforts [@refNikosinGermany] and also partially captured the hub ensembles propensity to forecast reductions in incidence after periods of increase. Unlike the forecast hub ensemble our surrogate model was well calibrated at all horizons and credible intervals explored though it was overly uncertain at short horizons. We found that the majority of the difference in performance between our surrogate model and the hub ensemble occured in the tails of the predictive distribution with similar central forecasts.

>>>>

  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 15th of August 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
  We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
  Data on notified weekly cases was originally sourced from the Johns Hopkins University curated data repository (JHU, [@REF]).
  We used the latest available observed data. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
  The follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of data revisions.
  Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.

@@

We extracted forecasts starting from the 15th of January until the 1st of August 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations). In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model). Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model. Of the remaining models submitted `r length(local_models)` were submitted in only one location. These models were gernerally submitted to a subset of locations with particular clustering in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project). Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else. `r nrow(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK. We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs and our observations of the performance of other models. We restricted ourselves to a model that could be easily understood, that produced meaningful summary statistics, and that could be run with low compute resources. We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.

Over the 6 months of the study period, we found that our surrogate model performed adequately at reproducing the forecast hub ensemble with average performance that was in general slightly worse though this varied by location and forecast date. Performance was closest for short forecast horizons and degraded relative to the forecast hub ensemble as the horizon increased. We observed that our surrogate model displayed similar levels of robustness to the forecast ensemble with few outlier forecasts unlike our previous forecasting efforts [@refNikosinGermany] and also partially captured the hub ensembles propensity to forecast reductions in incidence after periods of increase. Unlike the forecast hub ensemble our surrogate model was well calibrated for most horizons and credible intervals explored though it was overly uncertain at short horizons. We found that the majority of the difference in performance between our surrogate model and the hub ensemble occured in the tails of the predictive distribution with similar central forecasts.
